---
title: "Embedding choices (SCT vs LogNormalize) and the downstream pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Embedding choices (SCT vs LogNormalize) and the downstream pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE, comment = "#>")
library(TamuScDSC)
```

This vignette covers the two embedding knobs added to the integration helpers, and
how they feed the numbered pipeline scripts (DE, SCENIC, scTenifoldNet). The guiding
rule throughout: **the SCT assay is for the embedding/clustering/annotation only;
the `RNA` assay stays LogNormalized, and all differential expression reads `RNA`.**

## 1. Global embedding: `integrate_data()`

`integrate_data()` runs normalize → HVG → PCA → Harmony → clusters/UMAP and always
produces two tracks you can compare for batch effects: the un-integrated
`umap_unintegrated` / `clusters_unintegrated` and the Harmony `umap_harmony` /
`clusters_harmony`.

The normalization front-end is a toggle. `"LogNormalize"` (default) is the classic
path; `"SCT"` swaps in SCTransform v2 (glmGamPoi) before PCA.

```{r}
# LogNormalize (default)
obj <- integrate_data(obj, method = "RunHarmony", group_by = "SampleID")

# SCTransform embedding, optionally regressing a quality covariate
obj <- integrate_data(
  obj, method = "RunHarmony", group_by = "SampleID",
  normalization   = "SCT",
  vars_to_regress = "percent_mt"   # removes the dying/stressed-cell axis
)
```

Notes:

* SCTransform already models sequencing depth, so **never** put `nCount_RNA` in
  `vars_to_regress`. `percent_mt` is a legitimate quality covariate, but if mito
  fraction genuinely differs by condition/cell type it can erase real signal — the
  function prints a `[SCT] NOTE` when it regresses it. Compare with
  `normalization = "LogNormalize"` if unsure.
* `"SCT"` is wired to `method = "RunHarmony"` only.
* See `?integrate_data` for every parameter.

## 2. Subset re-embedding: `process_and_extract_cell_types()`

Pulls one or more cell types from any metadata column and re-embeds them. With
`harmonize_by` it emits **both** tracks (`umap_none`/`clusters_none` and
`umap_harmony`/`clusters_harmony`) so you can judge batch effects within the lineage;
`normalization = "SCT"` re-runs SCTransform on the subset (its HVGs differ from the
global object).

```{r}
sub <- process_and_extract_cell_types(
  obj, cell_types = "T cells", cell_type_col = "CellType_broad",
  normalization = "SCT", vars_to_regress = "percent_mt",
  harmonize_by  = "SampleID"
)
```

The subannotation scripts `03/04/05` call this via their `SUBCLUSTER_NORMALIZATION`
/ `SUBCLUSTER_VARS_TO_REGRESS` / `SUBCLUSTER_HARMONIZE_BY` config. See
`?process_and_extract_cell_types`.

## 3. Downstream pipeline scripts

These are standalone scripts (not package functions), documented in their header
comments and in `QUICK_START.md` / `INSTALL_NOTES.md`. They all read machine/project
paths from `config.R` (nr4a1 defaults; override via `NR4A1_*` env vars).

* **`07_DE_and_two_way_ANOVA.R`** — three DE tests side by side: **MAST**
  (sample-aware at cell level), **Wilcoxon** (`RUN_WILCOX`, sample-naive), and
  **edgeR pseudobulk** (`RUN_EDGER`, the sample-aware test for replicated designs),
  each with Enrichr and a DV∩test overlap. SplineDV adds differential-variability
  genes. Summary figures show all methods (collective grid + one per method; DV
  plotted unsigned).
* **`12a` → `12c` (SCENIC)** — `12a` exports per-label looms; `scenic.sh run`
  drives `run_pyscenic2.sh`/`12b` (grn→ctx→aucell); `12c` reads `aucell.csv` back,
  attaches an `AUC` assay, and makes FeaturePlot / violin / mean-AUC heatmaps.
* **`13_sctenifoldnet.R`** — differential gene-regulatory networks per cell type,
  with a resume-skip (`SKIP_IF_DONE`) so an interrupted run restarts cleanly.

## 4. Where to find help

```{r}
?integrate_data
?process_and_extract_cell_types
help(package = "TamuScDSC")
browseVignettes("TamuScDSC")
```
