---
title: "TamuScDSC: one pipeline, three starting points"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TamuScDSC: one pipeline, three starting points}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE, comment = "#>")
library(TamuScDSC)
```

The whole point of `TamuScDSC` is that *where your data comes from* does not change
*what you do to it*. `as_sample_list()` normalizes all three inputs to a named
list of per-sample Seurat objects; every step operates on that.

## Scenario 1 — one large merged Seurat object

```{r}
samples <- as_sample_list(big_object, sample_col = "SampleID")   # SplitObject
data    <- recipe_per_sample(samples)                            # or compose atoms
```

## Scenario 2 — per-sample checkpoint files (current lab layout)

```{r}
samples <- read_checkpoints("seurat_output/processed_files")     # reads *.rds
data    <- recipe_per_sample(samples, metadata = meta_df)
```

## Scenario 3 — a preloaded list of objects

```{r}
samples <- as_sample_list(my_list_of_seurats)                    # already a list
data    <- recipe_per_sample(samples)
```

## The two recipes (you vs. your colleague)

```{r}
# A: light per-sample QC, doublets BEFORE merge
data <- recipe_per_sample(input, metadata = meta_df)

# B: stringent merged QC first, doublets AFTER, then cluster review
data <- recipe_merged_stringent(input, metadata = meta_df)
```

## Composing atoms yourself (full control)

```{r}
data <- input |>
  as_sample_list() |>
  attach_metadata(meta_df) |>
  qc_light(qc_params(min_features = 200, max_mt = 20)) |>
  run_decontx() |>
  detect_doublets(method = "both", p = doublet_params(rate = "auto")) |>
  merge_samples() |>
  qc_stringent(qc_params(min_features_stringent = 500, max_mt_stringent = 10)) |>
  integrate_data(method = "RunHarmony")

# review clusters, then remove
data <- cluster_doublet_review(data, action = "flag", out_dir = "review")
data <- apply_doublet_action(data, action = "remove", target = "cluster_flag")
```

## Re-run one problem sample and graft it back

```{r}
sub  <- subset_samples(data, ids = "WT_3")
sub  <- detect_doublets(sub, method = "both", rate = 0.08)
data <- graft_meta(data, sub, cols = c("DF_score", "scDblFinder_score",
                                       "doublet_consensus"))
```

## Markers (used by annotation scripts 02–05)

```{r}
m <- get_markers()                                   # built-in default
m <- add_markers(m, "Tuft", c("Dclk1", "Trpm5", "Gfi1b"))
lst <- markers_as_list(m)                            # cell_type -> gene vector
```

## What happened to this object?

```{r}
provenance(data)
```
